Metabolite soup

### 2022-10-13 manuscript/manuscript.Rmd
Reformatting clockplot and adding metadata

Hi Regina

@@

```{r posthoc_classes and posthoc_table}
kw_diff_feats <- MT_peaks %>%
  left_join(filled_file_metadata, by="filename") %>% 
  select(feature, depth, value=norm_area) %>%
  nest(data=-feature) %>%
  mutate(tidyaov=map(data, ~ broom::tidy(kruskal.test(.x$value~.x$depth)))) %>%
  unnest(tidyaov) %>%
  mutate(pval=p.adjust(p.value, method = "fdr")) %>%
  filter(pval < 0.05) %>%
  select(feature, data) %>%
  unnest(data)
nondiff_feats <- MT_peaks %>%
  filter(!feature%in%kw_diff_feats$feature) %>%
  group_by(feature) %>%
  summarize(resp_class="All equal")

@@

```{r clockplot, fig.height=6, fig.width=6}
rank_peaks <- MT_peaks %>%
  group_by(feature) %>%
  mutate(norm_area=rank(norm_area)) %>%
  # mutate(value=value/max(value)) %>%
  # mutate(value=scale(value)[,1]) %>%
  ungroup()
clockplot_df <- rank_peaks %>%
  left_join(filled_file_metadata, by="filename") %>%
  select(feature, norm_area, depth, filename) %>%
  group_by(feature) %>%
  mutate(mean_area=mean(norm_area)) %>%
  group_by(feature, depth) %>%
  summarise(mean_fold_change=mean(norm_area/mean_area)) %>%
  pivot_wider(names_from=depth, values_from=mean_fold_change) %>%
  left_join(posthoc_classes, by="feature") %>%
  mutate(x=DCM-1) %>%       # cursed but works??
  mutate(y=`15m`+DCM/2-1.5) # super fucking cursed
clock_scale <- 0.66
axis_xends <- c(-1/2, 1.1, -1/2)*clock_scale
axis_yends <- c(sqrt(3)/2, 0, -sqrt(3)/2)*clock_scale
axis_y_dodge <- c(2, 2, -2)*clock_scale/25

@@

clock_gp <- ggplot(clockplot_df) +
  annotate("segment", x=c(0,0,0), y=c(0,0,0), 
           xend=axis_xends, yend=axis_yends,
           arrow=arrow(length=unit(0.02, "npc"), type = "closed")) +
  annotate("segment", x=c(0,0,0), y=c(0,0,0),
           xend=-axis_xends, yend=-axis_yends,
           lty = 2) +
  annotate("label", x=axis_xends, y=axis_yends+axis_y_dodge,
           label=c("15m axis", "DCM axis", "175m axis"), 
           fill="#FFFFFFAA", label.size=0) +
  geom_point(aes(x=x, y=y, color=resp_class), alpha=0.8) +
  coord_fixed(xlim = c(-1.1, 1.3)*clock_scale, ylim = c(-1.1, 1.1)*clock_scale) +
  scale_color_manual(values = color_scale, breaks = names(color_scale)) +
  theme_void() +
  theme(legend.position = "right", legend.title = element_blank())

@@

MAA_msms <- MAA_maybe %>%
  mutate(rtmed=rtmed/60) %>%
  pmap_dfr(function(...){
    row_data <- data.frame(...)
    msmsdata$MS2[rt%between%(row_data$rtmed+c(-1, 1))][premz%between%pmppm(row_data$mzmed, 50)] %>%
      mutate(MAA=row_data$MAA_name)
  }) %>%
  group_by(MAA, filename, rt) %>%
  arrange(desc(int)) %>%
  # slice(5) %>%
  summarize(MSMS=paste(fragmz, int, sep = ", ", collapse = "; "))

>>>>

@@

```{r posthoc_classes and posthoc_table}
long_MT_meta <- filled_file_metadata %>%
  filter(samp_type=="Smp") %>%
  filter(cruise=="MT") %>%
  select(filename, no23_um:euk_abund) %>%
  pivot_longer(cols = no23_um:euk_abund, names_to = "feature", values_to = "norm_area")
kw_diff_feats <- MT_peaks %>%
  bind_rows(long_MT_meta) %>%
  left_join(filled_file_metadata, by="filename") %>% 
  select(feature, depth, value=norm_area) %>%
  nest(data=-feature) %>%
  mutate(tidyaov=map(data, ~ broom::tidy(kruskal.test(.x$value~.x$depth)))) %>%
  unnest(tidyaov) %>%
  mutate(pval=p.adjust(p.value, method = "fdr")) %>%
  filter(pval < 0.05) %>%
  select(feature, data) %>%
  unnest(data)
nondiff_feats <- MT_peaks %>%
  filter(!feature%in%kw_diff_feats$feature) %>%
  group_by(feature) %>%
  summarize(resp_class="All equal")

@@

```{r clockplot, fig.height=6, fig.width=6}
rank_peaks <- MT_peaks %>%
  bind_rows(long_MT_meta) %>%
  group_by(feature) %>%
  mutate(norm_area=rank(norm_area)) %>%
  # mutate(value=value/max(value)) %>%
  # mutate(value=scale(value)[,1]) %>%
  ungroup()
clockplot_df <- rank_peaks %>%
  left_join(filled_file_metadata, by="filename") %>%
  select(feature, norm_area, depth, filename) %>%
  group_by(feature) %>%
  mutate(mean_area=mean(norm_area)) %>%
  group_by(feature, depth) %>%
  summarise(mean_fold_change=mean(norm_area/mean_area)) %>%
  pivot_wider(names_from=depth, values_from=mean_fold_change) %>%
  mutate(env_cmpd=ifelse(startsWith(feature, "MT"), FALSE, TRUE)) %>%
  left_join(posthoc_classes, by="feature") %>%
  mutate(x=DCM-1) %>%       # cursed but works??
  mutate(y=`15m`+DCM/2-1.5) # super fucking cursed
clock_scale <- 0.66
axis_xends <- c(-1/2, 1.1, -1/2)*clock_scale
axis_yends <- c(sqrt(3)/2, 0, -sqrt(3)/2)*clock_scale
axis_y_dodge <- c(2, 2, -2)*clock_scale/25

env_label_df <- clockplot_df %>%
  filter(!startsWith(feature, "MT")) %>%
  mutate(xend=x*1.3, yend=y*1.3) %>%
  mutate(xend=case_when(
    feature=="DON_um"~-0.5,
    feature=="syn_abund"~-0.3,
    feature=="DOP_um"~-0.1,
    feature=="pro_abund"~0.1,
    feature=="PC_um"~0.3,
    feature=="het_abund"~0.5,
    feature=="PN_um"~0.5,
    feature=="beam_atten"~0.7,
    TRUE~xend
  )) %>%
  mutate(yend=case_when(
    feature=="DON_um"~0.7,
    feature=="syn_abund"~0.8,
    feature=="DOP_um"~0.7,
    feature=="pro_abund"~0.8,
    feature=="PC_um"~0.7,
    feature=="het_abund"~0.8,
    feature=="PN_um"~0.6,
    feature=="beam_atten"~0.5,
    TRUE~yend
  ))

@@

clock_gp <- ggplot(clockplot_df) +
  annotate("segment", x=c(0,0,0), y=c(0,0,0), 
           xend=axis_xends, yend=axis_yends,
           arrow=arrow(length=unit(0.02, "npc"), type = "closed")) +
  annotate("segment", x=c(0,0,0), y=c(0,0,0),
           xend=-axis_xends, yend=-axis_yends,
           lty = 2) +
  annotate("label", x=axis_xends, y=axis_yends+axis_y_dodge,
           label=c("15m axis", "DCM axis", "175m axis"), 
           fill="#FFFFFFAA", label.size=0) +
  geom_point(aes(x=x, y=y, color=resp_class), alpha=0.8) +
  geom_segment(aes(x=x, y=y, xend=xend, yend=yend), data = env_label_df) +
  geom_label(aes(x=xend, y=yend, color=resp_class, label=feature), 
                   data = env_label_df, min.segment.length = 0) +
  # geom_label_repel(aes(x=x, y=y, color=resp_class, label=feature), 
  #                  data = env_label_df, min.segment.length = 0) +
  coord_fixed(xlim = c(-1.1, 1.3)*clock_scale, ylim = c(-1.2, 1.2)*clock_scale) +
  scale_color_manual(values = color_scale, breaks = names(color_scale)) +
  theme_void() +
  theme(legend.position = "right", legend.title = element_blank())

@@

MAA_msms <- MAA_maybe %>%
  mutate(rtmed=rtmed/60) %>%
  pmap_dfr(function(...){
    row_data <- data.frame(...)
    msmsdata$MS2[rt%between%(row_data$rtmed+c(-1, 1))][premz%between%pmppm(row_data$mzmed, 50)] %>%
      mutate(MAA=row_data$MAA_name)
  }) %>%
  group_by(MAA, filename, rt) %>%
  arrange(desc(int)) %>%
  # slice(5) %>%
  summarize(premz=mean(premz), MSMS=paste(fragmz, int, sep = ", ", collapse = "; "))
